---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
##load libraries
```{r libraries, echo=TRUE, message=FALSE}
library(dplyr)
library(Hmisc)
```
  
## Loading and preprocessing the data
```{r load, echo=TRUE}
activity <- read.table(unz("activity.zip", "activity.csv"), header = TRUE, sep = ",", quote = "\"")
totals <- activity %>% na.omit() %>% group_by(date) %>% dplyr::summarize(steps = sum(steps))
hist(totals$steps, breaks=53, main="Steps per day", xlab="Steps")
```
  
## What is mean total number of steps taken per day?
```{r mean, echo=TRUE}
m1 <- as.integer(mean(totals$steps))
```
###The mean is `r m1`.
```{r median, echo=TRUE}
m2 <- as.integer(median(totals$steps))
```
###The median is `r m2`.
  
## What is the average daily activity pattern?
```{r average, echo=TRUE}
fives <- activity %>% na.omit() %>% group_by(interval) %>% dplyr::summarize(steps = mean(steps))
with(fives, plot(x = interval, y = steps, type = "l"))
mx <- arrange(fives,desc(steps))[1,1]
```
###The interval with max steps is `r mx`.  
  
## Imputing missing values
```{r incomplete, echo=TRUE}
cc <- sum(!complete.cases(activity))
```
###There are `r cc` incomplete cases (rows with NA)
##I will use the impute function from the Hmisc package to fill in the NA values
```{r impute, echo=TRUE}
steps.i <- impute(activity$steps, mean)
activity.i <- activity
activity.i[,1] <- steps.i
totals.i <- activity.i %>% group_by(date) %>% dplyr::summarize(steps = sum(steps))
hist(totals.i$steps, breaks=53, main="Steps per day (imputed)", xlab="Steps")
m1.i <- as.integer(mean(totals.i$steps))
```
###The mean of imputed steps is `r m1.i`.
```{r median.i, echo=TRUE}
m2.i <- as.integer(median(totals.i$steps))
```
###The median of imputed steps is `r m2.i`.
  
###The median has increased by a single step.
###The addition of imputed data seems to have made no significant impact.
  
## Are there differences in activity patterns between weekdays and weekends?
```{r diffs, echo=TRUE}
vwdays <- as.factor(weekdays(as.Date(activity.i$date)))
wkdays <- as.factor(ifelse(vwdays %in% c("Monday","Tuesday","Wednesday","Thurdsday","Friday"), "weekday", "weekend"))
activity.i$wkday <- wkdays
fiveswkdays <- activity.i %>% filter(wkdays == "weekday") %>% group_by(interval) %>% dplyr::summarize(steps = mean(steps))
fiveswkends <- activity.i %>% filter(wkdays == "weekend") %>% group_by(interval) %>% dplyr::summarize(steps = mean(steps))
par(mfrow=c(2,1))
with(fiveswkdays, plot(x=interval, y=steps, type="l", main="Weekdays"))
with(fiveswkends, plot(x=interval, y=steps, type="l", main="Weekends"))

```
